\section{Introduction}
\label{sec:introduction}

During earthquake loading, saturated granular soils can experience a rapid rise in pore pressure that drastically reduces their shear strength. This phenomenon, termed liquefaction, has caused severe damage to infrastructure and buildings \citep{IshiharaKoga1981,SeedIdriss1967}. A large body of laboratory work has investigated liquefaction resistance using cyclic undrained triaxial tests, examining the effects of cyclic stress ratio (CSR), relative density, and confining pressure \citep{Hyodo1991,SeedLee1966,Silver1976,Toki1986}. However, vertically propagating shear waves in the ground induce gradually varying shear stress on soil elements, leading to continuous rotation of principal stress axes \citep{Arthur1980,IshiharaTowhata1983}, which cannot be fully represented in conventional triaxial loading.

Hollow cylinder apparatus (HCA) and hollow torsional shear tests address this limitation by allowing torsional loading with principal stress axis rotation. Early studies using HCA-type devices \citep[e.g.,][]{IshiharaYasuda1975,Tatsuoka1986,YamashitaToki1993} showed that liquefaction behavior can differ from triaxial-test results and depends on the loading mode and stress path. These observations motivate the use of HCA-type boundary conditions when investigating liquefaction mechanisms under more realistic cyclic loading paths.

The influence of the initial lateral-to-vertical effective stress ratio, \Kzero{}, on liquefaction resistance has been investigated extensively, yet the reported findings vary considerably depending on test conditions. Some experimental studies reported weak or negligible dependence of liquefaction strength on the initial anisotropic stress state \citep[e.g.,][]{IshiharaTakatsu1979,YamashitaToki1993}, whereas others found clear differences between isotropically consolidated (IC) and anisotropically consolidated (AC) specimens, with the trend depending on density \citep[e.g.,][]{GeorgiannouKonstadinou2014,Vargas2020}. More recently, \citet{Zhao2024} conducted systematic torsional shear tests covering both compressional and extensional consolidation states and attributed the higher liquefaction resistance under compressive consolidation to smaller principal stress axis rotation during cyclic loading. While these findings establish the macroscopic trend convincingly, the particle-scale mechanisms through which stress anisotropy alters liquefaction resistance remain to be explored.

The discrete element method \citep[DEM;][]{Cundall1979} is well suited for this problem because it provides direct access to particle-scale quantities such as contact topology, force chains, and fabric anisotropy, while avoiding some specimen-to-specimen variability associated with laboratory preparation. Numerous DEM studies have examined liquefaction-related behavior in triaxial, simple-shear, or polyhedral specimen settings \citep[e.g.,][]{Huang2018,Yang2021,Jiang2021,Morimoto2021,WuYang2022,Xie2023,YangHuang2023,Zhang2023} and have identified important microscopic factors governing liquefaction resistance. Recent DEM investigations have also addressed \Kzero{} effects: \citet{YangTaiebat2024} found that both preparation protocols and \Kzero{} influence liquefaction resistance, with the difference between IC and AC states narrowing at higher densities; \citet{Otsubo2022} showed that specimens with stronger inherent anisotropy exhibit weaker stiffness in the minor principal direction and correspondingly lower liquefaction resistance. However, triaxial configurations do not incorporate principal stress axis rotation, while periodic-boundary and polyhedral specimen settings, though capable of applying general stress paths, lack direct counterparts in standard laboratory apparatus. Although DEM replication of HCA tests has precedents under drained loading \citep[e.g.,][]{Li2014,Liu2021}, extension to undrained cyclic loading in HCA geometry remains limited. \citet{Han2024jgs} first realized undrained cyclic torsional shear in a DEM-based HCA model by introducing a combined servo mechanism that simultaneously satisfies the HCA stress boundary conditions and the constant-volume constraint. \citet{Ma2024} subsequently generalized this framework to accommodate a broader range of loading conditions, including drained stress paths, combined axial--torsional loading, and strain-controlled modes. Building on these developments, the present study applies the combined servo mechanism to systematically investigate \Kzero{} effects under undrained cyclic torsional shear with direct correspondence to laboratory HCA tests.

The present work investigates the effect of initial stress anisotropy on liquefaction resistance in DEM-simulated undrained cyclic torsional shear of hollow cylindrical specimens. Specimens are prepared using an isotropic-consolidation to anisotropic-consolidation (IC--AC) stress path, in which initial isotropic consolidation is followed by linear anisotropic consolidation from $\pmean=\SI{10}{kPa}$ to $\pmean=\SI{100}{kPa}$ to reach the target \Kzero{} state. The investigation is organized in two stages. First, the macroscopic \Kzero{}-dependent liquefaction resistance trend is established using three representative cases (\Kzero{} = 0.5, 1.0, and 2.0, corresponding to compression, isotropic, and extension states) at a dense state ($D_r \approx 90\%$) across multiple CSR levels, accompanied by macroscopic energy and stiffness analyses and microscopic fabric evolution. Second, to further disentangle the roles of packing compactness and directional fabric anisotropy, a medium-dense state ($D_r \approx 75\%$) is introduced alongside the dense state, and the \Kzero{} range is extended to five values (0.5, 0.67, 1.0, 1.5, and 2.0), yielding ten cases examined at CSR = 0.300. This two-stage approach first confirms the macroscopic \Kzero{} ordering and then demonstrates, through the expanded parametric set, that liquefaction resistance is jointly governed by macroscopic packing compactness and microscopic directional fabric anisotropy, with the coordination number and a signed fabric anisotropy indicator capturing complementary aspects of the observed trends.

The remainder of this paper is organized as follows. Section~\ref{sec:methodology} presents the DEM-HCA model, specimen preparation and loading protocol, calibration/validation strategy, and microscopic descriptors. Section~\ref{sec:results} reports the macroscopic and microscopic responses and discusses the mechanisms governing \Kzero{}-dependent liquefaction resistance. Section~\ref{sec:conclusions} summarizes the main findings and implications.
