\section{Conclusions}
\label{sec:conclusions}

This study used DEM simulations of undrained cyclic torsional shear in hollow cylindrical specimens to investigate how initial stress anisotropy (\Kzero{}) affects liquefaction resistance under HCA-type boundary conditions. Based on the results presented in this paper, the following conclusions are drawn.

For the representative dense-state simulations, liquefaction resistance exhibits a clear directional dependency on initial stress anisotropy. Compression-state anisotropy (\Kzero{} = 0.5) produces the highest resistance, the isotropic state (\Kzero{} = 1.0) shows intermediate resistance, and extension-state anisotropy (\Kzero{} = 2.0) yields the lowest resistance across the CSR range considered.

Macroscopic analyses indicate that the difference in liquefaction resistance cannot be explained solely by the total cumulative shear work required at liquefaction. While normalized energy accumulation patterns are similar among the three \Kzero{} states, the per-cycle energy input rate differs because stiffness degradation evolves differently. The compression-state specimen degrades stiffness more slowly and therefore accumulates shear work more gradually, whereas the extension-state specimen degrades stiffness faster and liquefies in fewer cycles.

Microscopic analyses show that initial coordination number and directional fabric anisotropy play distinct roles. The mechanical coordination number \Zm{} captures the effect of packing compactness and helps explain the difference between density levels, but it does not fully explain the \Kzero{} effect within a given density state. The signed fabric anisotropy indicator $\alpha$ and the corresponding contact-density morphology distinguish compression-type and extension-type anisotropy and correlate more directly with the variation in liquefaction resistance among \Kzero{} states.

Contact-density distributions and force-chain/displacement visualizations further show that initially distinct fabric morphologies progressively degrade and converge toward similar configurations near and after liquefaction. This convergence is consistent with the observed convergence in scalar fabric indicators and supports the interpretation that cyclic loading drives the granular assembly toward a common post-liquefaction fabric state despite different initial anisotropic stress conditions.

Overall, the results demonstrate that \Kzero{}-dependent liquefaction resistance in HCA-type cyclic torsional shear should be interpreted through both macroscopic and microscopic perspectives. The combination of DEM-based HCA simulation, energy/stiffness analysis, and fabric descriptors provides a practical framework for clarifying how initial stress anisotropy modifies liquefaction resistance in granular soils.
