\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{framed}
\usepackage{soul}
\usepackage{amsmath,amssymb}
\usepackage{siunitx}
\usepackage[authoryear,round]{natbib}
\usepackage{xr-hyper}
\usepackage{graphicx}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black,
}

% Pull labels (sections, figures, tables, equations) from the main
% manuscript's .aux file, so that \ref{sec:introduction} etc. resolve
% to the numbering in the compiled manuscript. Requires main.tex to
% have been built first.
\externaldocument{../../build/main}

% ---------------------------------------------------------------
% Response-letter environments and macros
% ---------------------------------------------------------------

% Reviewer's comment: left bar, italic, grey
\definecolor{reviewerbar}{gray}{0.55}
\definecolor{reviewerbg}{gray}{0.96}

\newenvironment{rcomment}
  {\begin{leftbar}\color{black!75}\itshape}
  {\end{leftbar}\vspace{0.3em}}

% Author's response: normal style, marked with a bold tag
\newcommand{\response}{\par\noindent\textbf{Response:}\ \ignorespaces}

% Manuscript change pointer
\newcommand{\changed}[1]{\par\noindent\textbf{Manuscript changes:}\ #1\par}

% Cross-reference helpers
\newcommand{\manref}[1]{(#1)}                  % e.g., \manref{p.~4, l.~112}
\newcommand{\secref}[1]{Section~#1}
\newcommand{\figref}[1]{Fig.~#1}
\newcommand{\eqnref}[1]{Eq.~#1}

% Commonly needed symbols in this paper
\newcommand{\Kzero}{\ensuremath{K_0}}
\newcommand{\Zm}{\ensuremath{Z_m}}
\newcommand{\Zmz}{\ensuremath{Z_{m0}}}
\newcommand{\Vs}{\ensuremath{V_s}}
\newcommand{\pmean}{\ensuremath{p^{\prime}}}
\newcommand{\ru}{\ensuremath{r_u}}

% ---------------------------------------------------------------

\title{Response to Reviewers\\
  \large Manuscript COMGE-D-26-01109}
\author{Yusong Han, Ryosuke Uzuoka, Kyohei Ueda}
\date{\today}

\begin{document}

\maketitle

\section*{Cover letter}

Dear Editor,

Thank you for the opportunity to revise our manuscript
``\emph{Influence of Initial Stress Anisotropy on Liquefaction Resistance
in Undrained Cyclic Torsional Shear: DEM Simulation of Hollow Cylinder
Tests}'' (COMGE-D-26-01109) submitted to
\emph{Computers and Geotechnics}. We sincerely thank both reviewers for their
thoughtful and constructive comments, which have helped us to substantially
improve the manuscript.

We have carefully addressed every comment. A point-by-point response is
provided below. In the revised manuscript, added or modified text is marked
using \texttt{latexdiff} against the submitted version (also provided).

%% -------- Summary of major changes --------

\section*{Summary of major revisions}

\begin{itemize}[leftmargin=*]
  \item \textbf{Clarified the relative-density labels}
    ``$D_r \approx 90\%$'' and ``$D_r \approx 75\%$'' as
    behaviour-matched references to the Toyoura density states
    reported by \citet{Nakata1998} and \citet{Ishihara1985}, rather
    than values computed from the DEM void ratio via the textbook
    formula.
  \item \textbf{Expanded the Introduction} with a review of prior DEM-HCA
    models and the boundary-condition taxonomy (rigid wall, stacked wall,
    flexible membrane).
  \item \textbf{Added a consolidated case table} listing all
    ($D_r$, $\Kzero$, CSR) combinations and the resulting $N_L$.
  \item \textbf{Added a quantitative quasi-static check} through the
    inertial number.
  \item \textbf{Added contact-force anisotropy analysis} to complement the
    existing contact-density analysis.
  \item \textbf{Quantified the differences} in Fig.~15 via ... .
  \item Minor clarifications and additional references throughout.
\end{itemize}

All line and figure numbers below refer to the revised manuscript unless
noted otherwise.

\clearpage

%% ===============================================================
%% Reviewer 1
%% ===============================================================

\section*{Response to Reviewer 1}

We thank Reviewer~1 for the detailed and constructive feedback. Our
responses to each point are given below.

%% ---------------- R1.1 ----------------
\subsection*{R1.1 --- Particle size scaling adequacy}

\begin{rcomment}
Lines 112--114: Particle size may influence macroscopic mechanical responses
such as dilatancy, peak friction angle, and liquefaction resistance. Please
clarify the adequacy of this scaling factor (e.g., whether the specimen
size-to-particle size ratio meets the requirement) or provide supporting
references.
\end{rcomment}

\response
We thank the reviewer for raising this important point. As the
reviewer notes, particle size choice can influence macroscopic
responses such as dilatancy, peak friction angle, and liquefaction
resistance. These three behaviours are precisely the targets against
which our calibrated parameter set has been validated (see item~(4)
below). The remainder of this response focuses on the more specific
question of whether the adopted specimen-to-particle size ratio is
adequate. We agree that the radial direction of an HCA specimen
(typically its thinnest dimension) is the direction most susceptible
to specimen-size effects and therefore deserves careful examination.
The present specimen is an annular hollow cylinder with inner
diameter \SI{60}{mm}, outer diameter \SI{100}{mm}, height
\SI{100}{mm}, and a particle-size distribution of
\SIrange{1.0}{3.0}{mm} ($d_{\max}=\SI{3}{mm}$). The corresponding
specimen-to-$d_{\max}$ ratios are 6.7 radially, 33 axially, and 84
in the mid-radius circumferential direction, confirming that the
radial ratio is the binding constraint. We examine its adequacy
through the following independent considerations.

\begin{enumerate}[leftmargin=*,label=(\arabic*)]
  \item \textbf{Codified laboratory guideline.} The standardised
    guidelines for triaxial testing both stipulate that the largest
    particle size shall be smaller than one-sixth of the specimen
    diameter: \citet{ASTM_D4767} for consolidated undrained triaxial
    testing on cohesive soils, and \citet{ASTM_D7181} for consolidated
    drained triaxial testing on soils generally (including cohesionless
    sands). The present radial ratio of 6.7 satisfies this 6:1
    threshold; the axial and circumferential ratios satisfy it by a
    wide margin.

  \item \textbf{DEM literature precedent.} Two strands of prior DEM
    work support the present size choice.
    \begin{enumerate}[label=(\alph*), leftmargin=2em, topsep=2pt, itemsep=3pt]
    \item \textit{Size-ratio effect on end-state and cyclic
      behaviour.} \citet{Huang2014} systematically investigated
      the effect of specimen/particle size ratio on DEM triaxial
      samples using a realistic Toyoura-sand grading and concluded
      that ``the ultimate strength of the material at the critical
      state is [\ldots] independent of the sample size and
      boundary conditions''; their smallest sample, at a
      diameter/$d_{\max}$ ratio of 7.2, is comparable to the
      present 6.7. In a more direct cyclic-liquefaction setting,
      \citet{KuhnRenken2014} simulated undrained cyclic shear in a
      DEM assembly with periodic boundaries and minimum cell
      dimension $\sim$12\,$D_{50}$ (comparable to the present
      radial/$D_{50}=10$), and reported that the resulting
      liquefaction resistance curves ($\mathrm{CSR}$ vs $N_L$) were
      nearly indistinguishable when the particle count was doubled.
      Because liquefaction ($\ru=0.95$) is itself an end-state
      condition characterised by vanishing effective stress, the
      quantity reported here ($N_L$) is expected to be weakly
      affected by specimen size in this regime.
    \item \textit{Comparable DEM-HCA configurations.} Within prior
      DEM-HCA studies, \citet{Song2024} simulated a DEM-HCA specimen
      with inner/outer diameters of 60/100~mm (identical to the
      present radial geometry) on upscaled Toyoura sand, using the
      same scaling strategy with an up-scaling factor somewhat
      larger than the 10$\times$ adopted here. \citet{Shi2021} and
      \citet{Liu2021} adopted radial/$d_{\max}$ ratios of 7.5 and
      7.2, respectively, in rigid-wall bonded-particle models of
      sandstone and mudstone under HCA torsional loading, both
      directly comparable to the present 6.7.
    \end{enumerate}

  \item \textbf{HCA-specific geometric argument.} Although the
    radial ratio of 6.7 is the tightest of the three, the primary
    shear strain $\gamma_{z\theta}$ under torsional loading
    develops in the axial--circumferential plane, \emph{orthogonal}
    to the thinnest (radial) dimension. The specimen/$d_{\max}$
    ratios in the two directions that carry this primary
    deformation are 33 (axial) and 84 (circumferential), both
    comfortably within the ranges established in~(1) and~(2a).

  \item \textbf{Agreement with laboratory data.} The calibrated parameter set
    reproduces both the monotonic undrained response of Toyoura sand
    reported by \citet{Nakata1998} (effective stress path, stress--strain
    relationship, and dilatancy) and the cyclic liquefaction-resistance
    curve of \citet{Ishihara1985}. Agreement across these two independent
    datasets, covering both monotonic and cyclic loading regimes,
    indicates that the adopted particle sizing captures the behaviours
    the model is used to predict.

\end{enumerate}

Taken together, these considerations support the adequacy of the
present size configuration for the quantities reported in this study.

\changed{In the revised manuscript we have (i) added citations to
\citet{Shi2021} and \citet{Song2024} in the Introduction's DEM
paragraph (Section~\ref{sec:introduction}) in support of this
response, and (ii) added a brief note in
Section~\ref{subsec:dem_hca} citing \citet{Song2024} as precedent
for the 10$\times$ scaling of the Toyoura-sand grading.}

%% ---------------- R1.2 ----------------
\subsection*{R1.2 --- Void ratio range and ``dense''/``medium-dense'' labels}

\begin{rcomment}
For the case where the authors have defined relative density states of
$D_r \approx 90\%$ and $D_r \approx 75\%$ based on void ratios of 0.736 and
0.742, respectively, the implied range of maximum and minimum void ratios is
approximately 0.732 to 0.772. As a result, I do not think that a void ratio
range of merely 0.04 and the terms ``dense'' and ``medium-dense'' as used in
this manuscript correspond to the physical states observed in the Toyoura
sand that this model purports to represent.
\end{rcomment}

\response
We thank the reviewer for this careful observation. The reviewer's
arithmetic correctly identifies that reading
``$D_r \approx 90\%$'' and ``$D_r \approx 75\%$'' as strict point
values on Toyoura's tabulated $(e_{\max}, e_{\min})$ scale would
imply an implausibly narrow void-ratio range. The underlying issue,
we believe, is that \emph{relative density is not a well-defined
quantity for a DEM assembly}, and the labels in our manuscript are
behaviour-matched references to laboratory conditions rather than
quantities computed from the DEM void ratio.

\medskip
\noindent\textbf{Why $D_r$ does not transfer from laboratory
Toyoura to this DEM assembly.}
The particles in the present model are spherical, whereas real
Toyoura sand consists of subrounded-to-subangular grains; packing
under sedimentation or shearing depends strongly on particle shape,
so the achievable void ratio of a spherical-particle assembly
differs systematically from that of an angular-particle assembly of
the same nominal grading. Furthermore, no standardised procedure
exists for determining the loosest and densest states of a DEM
assembly: the attainable $e_{\max}$ and $e_{\min}$ depend on the
choice of particle shape and grading, the inter-particle friction
used during assembly, the numerical scheme and damping, and the
boundary and preparation conditions, none of which are standardised
across DEM codes. Relative density, strictly speaking, is therefore
not a meaningful quantity in a DEM context in the way it is for
laboratory sand. Consistent with this, substituting the DEM void
ratios (0.736, 0.742) into the textbook $D_r$ formula with
real-Toyoura limits $(e_{\max}, e_{\min})=(0.98, 0.60)$ reported by
\citet{Ishihara1985} yields $D_r \approx 64\%$ for both
specimens---inconsistent with the cyclic behaviour they actually
reproduce---which confirms that the textbook formula does not apply
to this DEM assembly.

\medskip
\noindent\textbf{What the labels ``$D_r \approx 90\%$'' and
``$D_r \approx 75\%$'' actually represent.}
Rather than computing $D_r$ from the DEM void ratio, we tuned the
target void ratio so that the resulting DEM specimen reproduces the
macroscopic response of real Toyoura sand at the corresponding
$D_r$ condition as reported in the laboratory literature.
Specifically:
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=3pt]
  \item The ``$D_r \approx 90\%$'' DEM specimen ($e \approx 0.736$)
    simultaneously reproduces (a) the undrained monotonic stress
    path, stress--strain response, and dilatancy behaviour of
    \citet{Nakata1998}, whose Toyoura specimens were prepared at
    $D_r \approx 92\!-\!93\%$, and (b) the cyclic
    liquefaction-resistance curve of \citet{Ishihara1985} for their
    ``very dense'' Toyoura band ($D_r = 88\!-\!95\%$) at
    $K_0 = 1.0$. Both comparisons are shown in
    Fig.~\ref{fig:validation_combined} of the revised manuscript.
  \item The ``$D_r \approx 75\%$'' DEM specimen ($e \approx 0.742$)
    serves as a comparative density anchor for the $K_0$-expansion
    study, chosen so that its macroscopic cyclic response at
    $K_0 = 1.0$ falls within the ``dense'' ($D_r = 75\!-\!82\%$)
    cyclic-resistance data of \citet{Ishihara1985} at representative
    loading amplitude.
\end{itemize}

\medskip
\noindent\textbf{Precedent for band-based labelling.}
\citet{Ishihara1985} himself groups test specimens into two
density bands rather than point values---``very dense''
($D_r = 88\!-\!95\%$) and ``dense'' ($D_r = 75\!-\!82\%$), each
$\sim$7\%-wide. The paper notes explicitly that
``\emph{individual specimen has a variation in relative density
with a range of $\pm 3.5\%$ around the average}'' (p.~68), and that
the specimen height had to be restored manually at the end of each
cycle of undrained torsional loading because the specimen cannot
automatically recover its deformation between half-cycles (p.~67).
The ``$D_r \approx 90\%$'' and ``$D_r \approx 75\%$'' labels in our
manuscript follow the same band-based convention.

\changed{Section~\ref{subsec:preparation}: revised the paragraph
introducing the two DEM density states to clarify (i) that the DEM
void ratios cannot be mapped onto Toyoura's textbook relative
density, because of the spherical particle shape and the absence of
a standardised DEM packing-limit procedure; and (ii) that the
labels ``$D_r \approx 90\%$'' and ``$D_r \approx 75\%$'' are used
as behaviour-matched references to Ishihara's ``very dense''
($D_r = 88\!-\!95\%$) and ``dense'' ($D_r = 75\!-\!82\%$) density
bands, tuned to reproduce the corresponding macroscopic monotonic
and cyclic responses. No figure changes in the main manuscript.}

%% ---------------- R1.3 ----------------
\subsection*{R1.3 --- Missing formulas for CSR and $\Kzero$}

\begin{rcomment}
Several variable symbols are used without a clear computational formula.
Such as CSR, $\Kzero$.
\end{rcomment}

\response
Both ratios are now defined inline where each symbol first appears
in Section~\ref{subsec:preparation}. $\Kzero$ is introduced at the
opening of the IC--AC consolidation protocol as the ratio of radial
to vertical effective stresses,
$\Kzero = \sigma_r^{\prime}/\sigma_z^{\prime}$; and CSR is introduced
at the opening of the cyclic shear program as the amplitude of
cyclic shear stress normalised by the initial mean effective stress,
$\mathrm{CSR} = \tau_{cyc}/p_0^{\prime}$. Both definitions
cross-reference Table~\ref{tab:hca_equations}, which collects the
full set of HCA stress and strain expressions.

\changed{Inline definitions of $\Kzero$ and $\mathrm{CSR}$ added in
Section~\ref{subsec:preparation} (IC--AC protocol opening and
cyclic-shear program opening, respectively), with cross-references
to Table~\ref{tab:hca_equations}.}

%% ---------------- R1.4 ----------------
\subsection*{R1.4 --- Consolidated case table}

\begin{rcomment}
The description of the simulation program is summarized in brief statements
without explicitly listing the complete set of parameters. It is recommended
that all simulation cases (including combinations of different densities,
$\Kzero$ values, and CSR levels) be presented collectively in a table
format.
\end{rcomment}

\response
A consolidated table has been added at the end of
Section~\ref{subsec:preparation} (Table~\ref{tab:case_matrix})
listing all simulation cases. Each row identifies a
$(D_r, \Kzero)$ combination; the third column enumerates the CSR
levels tested for that case. The primary set at $D_r \approx 90\%$
covers $\Kzero \in \{0.5, 1.0, 2.0\}$ across five CSR levels
($0.200$ to $0.400$ in $0.05$ steps); the expanded set adds two
intermediate $\Kzero$ values ($0.67$ and $1.5$) at $D_r \approx
90\%$ together with the full $D_r \approx 75\%$ strip, all at
CSR $= 0.300$. The resulting $N_L$ values are presented in the
Results section (e.g., Fig.~\ref{fig:liq_resistance_curves}) so
that the methodology table itself remains free of outcome data.

\changed{New Table~\ref{tab:case_matrix} inserted at the end of
Section~\ref{subsec:preparation}, summarising the full simulation
program with CSR levels enumerated for each $(D_r, \Kzero)$ case.}

%% ---------------- R1.5 ----------------
\subsection*{R1.5 --- Loading rate / quasi-static condition}

\begin{rcomment}
The loading rate is not specified in the text. In DEM simulations, an
excessively high loading rate may introduce inertial effects, leading to
excessive particle accelerations and distorted contact forces. Please
clarify whether the selected loading rate satisfies quasi-static conditions.
\end{rcomment}

\response
\textit{[TODO: report inertial number $I = \dot{\gamma}\,d\,\sqrt{\rho/p}$;
show $I \ll 10^{-3}$; cite da Cruz / Roux.]}

\changed{\textit{[TODO: \secref{2.3}, new equation / paragraph.]}}

%% ---------------- R1.6 ----------------
\subsection*{R1.6 --- Rigid wall vs.\ flexible membrane}

\begin{rcomment}
Rigid walls are used to model the inner and outer boundaries of the HCA
specimen, whereas flexible latex membranes are typically used in laboratory
experiments. Rigid boundaries may restrict local deformation and shear band
development, potentially leading to overestimation of stiffness and
liquefaction resistance. Please discuss the limitations of adopting rigid
boundaries and justify their use with relevant references.
\end{rcomment}

\response
\textit{[TODO: acknowledge limitation; justify via (i) HCA chambers are
rigid-walled with the membrane as a pressure interface, (ii) present work
focuses on average fabric-level comparison rather than localized shear
bands, (iii) flexible-membrane DEM-HCA (Song 2024) as complementary / future
work.]}

\changed{\textit{[TODO: \secref{2.1} end paragraph or discussion section.]}}

%% ---------------- R1.7 ----------------
\subsection*{R1.7 --- Citation for CSL in Fig.~8(c)}

\begin{rcomment}
The CSL in Fig.~8(c) is of unknown origin; whether it comes from your own
data or from someone else's literature, a clear citation is required.
\end{rcomment}

\response
\textit{[TODO: verify source; update caption with explicit citation.]}

\changed{\textit{[TODO: \figref{8}(c) caption.]}}

%% ---------------- R1.8 ----------------
\subsection*{R1.8 --- ``Enhancement'' of $\Kzero = 1.0$ in the coordination-number plane (Fig.~12)}

\begin{rcomment}
In most cases, $N_L$ decreases monotonically with the increase of $\Kzero$,
as shown in Figures 8 and 10; however, the situation changes in Figure 12.
Theoretically, the evolution of the coordination number at the microscopic
level corresponds to the evolution of the macroscopic parameters such as the
excess pore water pressure ratio, deviatoric stress, and cumulative shear
work. The enhancement of $\Kzero = 1.0$ in the coordination number plane
should be explained more clearly.
\end{rcomment}

\response
\textit{[TODO: clarify physical cause --- isotropic consolidation distributes
contact normals more uniformly, yielding slightly higher $\Zmz$; reaffirm
that $\Zmz$ alone is insufficient to predict $N_L$ and that the directional
fabric indicator $\alpha$ is required, as demonstrated in
\secref{...}.]}

\changed{\textit{[TODO: 1--2 sentences appended to \figref{12} discussion.]}}

\clearpage

%% ===============================================================
%% Reviewer 2
%% ===============================================================

\section*{Response to Reviewer 2}

We thank Reviewer~2 for the thorough and constructive comments, and
particularly for pointing us to relevant prior work on DEM-HCA modelling.

%% ---------------- R2.1 ----------------
\subsection*{R2.1 --- Particle--wall friction coefficient set to zero (Table 1)}

\begin{rcomment}
Table 1: Why is the friction coefficient between particles and walls zero?
\end{rcomment}

\response
The zero particle--wall friction coefficient applies uniformly to all
wall-type elements in the PFC model --- the inner and outer
cylinders, the top and bottom platens, and the six torsional blades
(which are also wall elements in PFC). The physical justification
differs by wall class, and we organise the response accordingly.

\medskip
\noindent\textbf{Cylindrical walls --- radial principal stress plane.}
The HCA element-test reduction used in this work (Appendix~A,
Table~\ref{tab:hca_equations}) assumes the radial-normal face to be a
principal stress plane, which requires zero shear traction on it. In
a laboratory HCA, the specimen is wrapped by a flexible membrane
immersed in a pressurised water chamber. The membrane does carry
non-negligible sand--membrane friction at each point of contact, but
under small-strain quasi-static loading the membrane is in mechanical
equilibrium: the local tangential tractions it transmits to the sand
balance around the circumference and integrate to zero at the
element scale. The radial-normal face therefore remains a principal
stress face in the laboratory. In the DEM model, by contrast, the
wall elements are massless and cannot carry their own equilibrium;
prescribing $\mu_{pw}>0$ would leave a residual shear resultant on
the radial-normal face, violating the principal-plane assumption on
which the HCA stress reduction rests. Setting $\mu_{pw}=0$ on the
cylinders is therefore \emph{required} (not merely a simplification)
to preserve the element-test idealisation that the laboratory
apparatus delivers through its membrane--water balance.

\medskip
\noindent\textbf{End platens --- axial principal stress plane and
servo determinacy.}
The same principal-plane argument applies to the axial-normal face.
Laboratory HCA end platens (porous-stone caps) do carry some residual
sand--cap friction, a well-known experimental non-ideality that
contaminates the axial principal-stress state. The present DEM model
represents the idealised HCA boundary condition rather than this
laboratory non-ideality: $\mu_{pw}=0$ on the end platens keeps the
$\sigma_z^{\prime}$ face as a principal stress face, and additionally
prevents a Coulomb-limit coupling between the vertical stress and the
torsional loading that would otherwise be introduced by a nonzero cap
friction and that cannot be constrained by any measurable laboratory
quantity.

\medskip
\noindent\textbf{Blades --- geometric torque transmission.}
Torque is transmitted from the end platens into the specimen through
the six torsional blades. Each blade is a thin radial slab whose
face normal points in the circumferential direction, so a normal
contact between a particle and a blade face delivers a purely
circumferential force --- the torque-transmitting direction. Full
torque transfer is therefore guaranteed geometrically and does not
require tangential friction at the blade--particle interface.
Conversely, prescribing $\mu_{pw}>0$ at the blade surface would
introduce Coulomb-limited frictional tractions in the blade's
tangent plane (the radial--axial plane). These tractions would
(i)~directly contaminate the axial stress measurement, since
$p_z^{\prime}$ (Appendix~A) is integrated from the end-platen contact
forces including those on the attached blades, and (ii)~introduce a
Coulomb-dependent contribution to the blade--particle force
distribution that saturates as $\sigma_r^{\prime}\to 0$ in the
post-liquefaction regime, rendering the servo response non-smooth
and causing it to miss the prescribed $\tau^{tar}$.

\medskip
Taken together, $\mu_{pw}=0$ (i)~preserves the element-test
principal-stress assumption on both the radial- and axial-normal
faces, (ii)~avoids a Coulomb-limit servo constraint that would
compromise torque control near liquefaction, and (iii)~removes a
free parameter unconstrained by any measurable laboratory quantity.

\changed{Section~\ref{subsec:dem_hca}: added a short clause after the
torsional-blade paragraph clarifying that the cylindrical walls and
end platens are assigned $\mu_{pw}=0$ so that they act as pure
pressure boundaries, with torque transmission provided geometrically
by the blade faces. The full principal-plane and servo-determinacy
argument is kept in this letter rather than added to the manuscript,
since it concerns a design-rationale question orthogonal to the
paper's focus on $K_0$ effects and fabric evolution.}

%% ---------------- R2.2 ----------------
\subsection*{R2.2 --- Size of torsional blades}

\begin{rcomment}
What is the size of torsional blades?
\end{rcomment}

\response
The blade height is \SI{15}{mm}, and the blades extend radially
across the full annulus between the inner and outer cylinders. Six
blades are inserted from the top and bottom plates (three each),
uniformly spaced in the circumferential direction; their arrangement
is visible in Fig.~\ref{fig:specimen_generation}.

The blades of the laboratory HCA apparatus are approximately
\SI{1.5}{mm}, which is \SI{15}{} times the minimum particle size of
Toyoura sand. In the present DEM model, the blade height is
correspondingly set to \SI{15}{} times the minimum particle size of
the DEM grading ($d_{\min}=\SI{1.0}{mm}$, see the
\SIrange{1.0}{3.0}{mm} grading stated in
Section~\ref{subsec:dem_hca}), giving \SI{15}{mm}. This preserves
the blade-to-grain scaling of the laboratory apparatus.

\changed{Section~\ref{subsec:dem_hca}: the blade-insertion sentence
now states the blade height (\SI{15}{mm}) explicitly, followed by a
short sentence giving the blade layout (three blades at each end of
the specimen, extending radially across the annulus and uniformly
spaced around the specimen axis). The blade-to-grain-size rationale
is kept in this letter rather than added to the manuscript: the
methodology section reports the blade height as a model parameter,
but the paper does not otherwise discuss whether the
blade-to-$d_{\min}$ ratio is appropriate for DEM-HCA modelling,
which is orthogonal to the paper's focus on $K_0$ effects and
fabric evolution.}

%% ---------------- R2.3 ----------------
\subsection*{R2.3 --- Difference vs.\ servo mechanism in Han (2024)}

\begin{rcomment}
What is the difference between the servo mechanism used in this paper and
that in Han (2024)?
\end{rcomment}

\response
The contribution paragraph at the end of Section~\ref{subsec:cyclic_loading}
(already in the submitted manuscript) enumerates three explicit
differences from \citet{Han2024jgs}, which we summarise here for
convenience.
\begin{enumerate}[leftmargin=*,label=(\roman*),topsep=2pt,itemsep=3pt]
  \item \textbf{Degrees of freedom.} \citet{Han2024jgs} solved only
    the radial and torsional conditions, with the specimen height
    held fixed ($dH/dt=0$); axial loading and drained/undrained
    height variation were outside the scope. The present
    formulation adds the axial degree of freedom as a first-class
    unknown, enabling general undrained and drained loading paths
    with coupled radial, axial, and torsional components.
  \item \textbf{Laboratory-to-DEM mapping.} \citet{Han2024jgs}
    targeted individual effective stresses directly, which is only
    an approximation because the pore water pressure $u$ is a
    measured response, not a prescribed input. The present servo
    instead targets the two pressure differences $p_{dif,r}^{\prime}$
    and $p_{dif,z}^{\prime}$ (Section~\ref{subsec:cyclic_loading}), which are
    invariant to $u$ and therefore provide a mapping to the
    laboratory total-stress boundary conditions that does not
    require $u$ as input.
  \item \textbf{Cross-coupling.} In \citet{Han2024jgs}, with
    $dH/dt$ suppressed, the radial condition reduces to a scalar
    and cross-coupling does not arise. In the present formulation,
    once $dH/dt$ is retained, $dr/dt$ and $dH/dt$ jointly affect
    both the radial and axial stresses, and the servo coefficients
    must reflect this coupling. We derive the $2\times 2$ coupled
    stiffness matrix $\hat{\mathbf{K}}$ analytically and invert it
    in closed form at every timestep, so the off-diagonal entries
    $S_{rz}$ and $S_{zr}$ correctly represent the mutual influence
    of $dr/dt$ and $dH/dt$ on both boundary conditions.
\end{enumerate}
Under the constant-height condition ($dH/dt=0$) adopted by
\citet{Han2024jgs}, the present formulation reduces exactly to their
radial-only servo coefficient, as stated after
Eq.~(\ref{eq:servo_dr})--(\ref{eq:servo_dH}) in
Section~\ref{subsec:cyclic_loading}; in that limiting case, the present and
prior formulations are equivalent.

\changed{No additional text has been added on this point. The
existing three-contribution paragraph at the end of
Section~\ref{subsec:cyclic_loading} already identifies each of these
differences, and the reduction to the \citet{Han2024jgs} formulation
under $dH/dt=0$ is shown there as well.}

%% ---------------- R2.4 ----------------
\subsection*{R2.4 --- DEM-HCA literature review and the $H=D$ height choice}

\begin{rcomment}
There should be more context to introduce the research in DEM HCA models in
Part Introduction. \dots\ [boundary-condition types in Li~2014, Li~2017,
Song~2024a, Song~2024b; height $H \approx 2D$ in Shi~2021]. However, the
height equals to the outer diameter in this paper. Are there any
requirements or guidelines for the size of HCA?
\end{rcomment}

\response
We thank the reviewer for pointing us to this body of prior DEM-HCA
work. Our response has two parts: an expanded introduction of prior
DEM-HCA research, and a justification of the $H=D$ aspect ratio.

\medskip
\noindent\textbf{Prior DEM-HCA research and boundary taxonomy.}
The submitted manuscript has been revised to cite the five DEM-HCA
studies identified by the reviewer, grouped by the radial
boundary-condition treatment that distinguishes them. Within
rigid-boundary or stacked-wall formulations,
\citet{LiGuoZhang2014} applied cyclic torsional loading under
drained conditions using a stacked-wall boundary to mimic membrane
compliance, and \citet{LiZhangGutierrez2015} used the same family
of boundaries to analyse principal-stress-axis rotation in drained
monotonic loading; \citet{Shi2021} and \citet{Liu2021} adopted
rigid cylindrical walls combined with bonded-particle assemblies
to analyse rock-like HCA behaviour. A more recent line of work by
\citet{Song2024} introduced a flexible bonded-sphere membrane as
the radial boundary. The present formulation falls within the
rigid-wall class and extends it to undrained cyclic torsional
loading. These citations now appear in the DEM paragraph of
Section~\ref{sec:introduction}.

\medskip
\noindent\textbf{Specimen aspect ratio ($H=D_{\text{outer}}$).}
Laboratory HCA specimens typically adopt $H/D_{\text{outer}}$ in
the range $\sim\!1.0$--$2.0$. The longer aspect ratios are
preferred for studies that target strain localisation and
shear-band formation, where a longer specimen reduces end-platen
constraint on the band trajectory. The present work, in contrast,
targets the \emph{average specimen-level} response: the
macroscopic liquefaction resistance curves
(Fig.~\ref{fig:liq_resistance_curves}) and the
specimen-averaged fabric quantities
($Z_{m0}$, $\alpha$, $\rho_c$) in
Section~\ref{sec:results}. All stress and strain expressions used
in the analysis (Table~\ref{tab:hca_equations}) are integrated
over the entire specimen boundary, and the fabric descriptors are
evaluated over the full contact population; the analysis therefore
does not rely on spatial resolution of localised deformation.
For this class of problem, aspect-ratio requirements are
considerably relaxed, and the $H=D$ configuration matches the
laboratory-scale HCA apparatus cited in Section~\ref{subsec:dem_hca}
\citep{Vargas2020}; the same geometry was also adopted by the
DEM-HCA servo-formulation precursors \citet{Han2024jgs} and
\citet{Ma2024}. Choosing $H=D$ rather than $H=2D$ approximately
halves the particle count at fixed grading, which is a material
saving given the parametric scope of this study
(Table~\ref{tab:case_matrix}) and the per-case cost of high-cycle
undrained simulations near liquefaction.

\changed{Section~\ref{sec:introduction}, DEM paragraph: expanded to
group the cited DEM-HCA precedents by radial boundary treatment ---
rigid cylindrical walls \citep{Shi2021,Liu2021}, stacked-wall
approximations of membrane compliance
\citep{LiGuoZhang2014,LiZhangGutierrez2015}, and flexible
bonded-sphere membranes \citep{Song2024} --- and to position the
present work as an extension of the rigid-wall class to undrained
cyclic loading. No dedicated text on the aspect-ratio choice was
added to the manuscript; the justification above is provided in the
reviewer's frame without expanding Section~\ref{subsec:dem_hca}.}

%% ---------------- R2.5 ----------------
\subsection*{R2.5 --- Stability of the servo near singular stiffness matrix}

\begin{rcomment}
The proposed servo mechanism may face numerical instability when the
stiffness matrix becomes near-singular due to the drastic loss of particle
contacts. How does the algorithm prevent numerical oscillations during
matrix inversion?
\end{rcomment}

\response
\textit{[TODO: depending on code audit --- either (a) describe
regularisation / condition-number checks, or (b) argue that the
constant-volume constraint eliminates radial-axis degeneracy, while the
torsional and axial degrees of freedom remain independently stable; any
post-liquefaction oscillations observed are physical rather than numerical.]}

\changed{\textit{[TODO: paragraph in \secref{...} or appendix.]}}

%% ---------------- R2.6 ----------------
\subsection*{R2.6 --- Evolution of $e_r$, $e_z$, or volume during validation}

\begin{rcomment}
For the validation part, please add some figures to display the evolution
of $e_r$ and $e_z$ over time or the change in volume.
\end{rcomment}

\response
\textit{[TODO: add plots of $r$, $H$, and total volume during monotonic
calibration and cyclic validation, confirming strict enforcement of the
constant-volume constraint.]}

\changed{\textit{[TODO: new panel in \figref{7} or a separate
\figref{...}.]}}

%% ---------------- R2.7 ----------------
\subsection*{R2.7 --- Contact-force anisotropy (3D rose diagrams)}

\begin{rcomment}
Fig.~14: You analysed contact density anisotropy, but why didn't you analyse
anisotropy for contact forces such as normal contact force? Please refer to
the papers mentioned before which plotted 3D rose diagrams of contact force
anisotropies.
\end{rcomment}

\response
\textit{[TODO: add contact-force rose diagrams alongside the contact-density
ones; reference Song~2024 precedent.]}

\changed{\textit{[TODO: extended \figref{14} or new companion figure.]}}

%% ---------------- R2.8 ----------------
\subsection*{R2.8 --- Quantifying the differences in Fig.~15}

\begin{rcomment}
Fig.~15: It is difficult to distinguish the difference in displacement
between $\Kzero=0.5$ and $2.0$ specimens when $N/N_L=0.00$ and $N/N_L=0.61$,
and it is also difficult to distinguish the difference in force chain when
$N/N_L=1.05$ and $N/N_L=1.07$. Please quantify the differences.
\end{rcomment}

\response
\textit{[TODO: add a quantitative metric --- e.g., displacement-magnitude
CDF and force-chain principal-direction distribution --- either as a new
panel or in the caption.]}

\changed{\textit{[TODO: \figref{15} caption / new panel.]}}

\clearpage

\section*{Closing}

We hope the revised manuscript now meets the journal's standards and the
reviewers' expectations. We remain grateful for the reviewers' careful
reading and valuable suggestions.

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}
